Experience shapes individual foraging specialization and success in a virtual predator-prey system

### 2023-10-20 manuscript/Chapter2-manuscript.Rmd
updated refs and intro. getting there

Predator express a diverse array of hunting tactics to capture their prey. In natural populations, individual differences in tactics such as slow and fast movement speeds or high attack frequencies are associated with differences in prey choice and prey consumption [@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@toscanoPersonalityForagingBehavior2016]. Prey usually respond to predation cues by reducing their activity, or use rapid or unpredictable movements to successfully escape when they are chased [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Such antipredator tactics often result in reciprocal interactions where predators fine-tune their tactics to the type of prey that they encounter [@Estes.etal2003;@Woo.etal2008;@Phillips.etal2017;@Heithaus.etal2018;@Courbin.etal2018]. This has prompted ecologists to hypothesize that individual predators develop their hunting tactics by learning through repeated interactions with their prey. Thus, predators may only be successful if they learn how to hunt their prey by refining their hunting tactics, which should require periods of extensive practice, i.e., experience [@DeLisle.etal2022;@Tinker.etal2009;@Dukas2019].

Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. As predators practice and refine their tactics, they acquire expertise (e.g. skills, knowledge) that allows them to outperform novice hunters [@Dukas2019]. Through this process, it is hypothesized that individuals may specialize in a hunting tactic (e.g. always moving fast) if it allows them to maintain successful prey captures at each atempt [@Tinker.etal2009;@Dukas2019]. An alternative view is that learning operates as a mechanism driving behavioural flexibility. For instance, predators can gain expertise and information on their prey through repeated hunting atempts, and may thus hunt optimally by adjusting their behaviour to the type of prey that they encounter [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013].

Ecological interactions when individuals are novice can also shape learning and behavioural trajectories [@vilapouca2021;@beyts2023], making it challenging to predict under which circumstances individuals should specialize in their foraging with experience (REFS).

**Parler des risques pour la transition et les couts/bénéfices qui varient selon le contexte**
(ici, je dois toutefois faire ressortir qu'il est dangeureux d'être spécialiste si les ressources fluctuent beaucoup.)
Moreover, individual specialists may avoid the risks of loosing a meal by testing other tactics through trial and error, which can be highly costly if prey are scarce.

Alternatively,  the costs/risks faced by hunters when switching tactics may be offset through extensive practice,

To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]

However, predators can differ in the challenges that they face during their lifetime, which could result in both processes (i.e. specialization and flexibility) operating at the same time within a predator population. For example, some individuals may hunt in habitats where prey are faster, which are often proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022].

*dans mcghee :*
*The coexistence of a range of consistent behavioral types within prey (or predator) populations (Sih et al. 2004a, 2004c) creates a situation where individual prey (or predators) are not equivalent to one another in terms of predation risk (or threat).*

Deux idées pour le paragraphe suivant :
1. Focus sur les contextes qui génèrent des différences, et qui devraient donc aussi mener à des différences de fitness (moins mettre l'accent sur le manque de consensus et plutôt focus sur les contextes, pour ensuite expliquer que donc ça devrait être mieux dans un que l'autre)
2. Je pourrais aussi commencer avec l'exemple indiquant que le foraging behaviour est lié aux proies rencontrées ou capturées

There is currently a lack of consensus on the fitness advantages of specialized vs flexible hunting strategies [@Phillips.etal2017]. Some studies have found greater success for individuals specializing on a hunting tactic [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others have found that individuals using flexible hunting tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021]. This discrepancy may in part be due to the limited information we have on the ecological contexts favoring individual specialization over flexibility. However,

(*pourrait retirer en haut et commencer ici*)
many studies also show that both type of foragers can attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018]. It is hypothesized that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit from adjusting to fluctuating resources by reducing the consequences of uncertainty [@dall2010]. While it is clear that predators can match their hunting behaviour and locomotion to their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022], the important question remaining is how. Given the dynamic nature of predator-prey interactions, it is likely that ontogenic behavioural changes in predators play an important role, because complex hunting strategies require practice to be effective [@Kelley.Magurran2011].

>>>>

Predator express a diverse array of hunting tactics to capture their prey. In natural populations, differences among individuals in tactics such as slow and fast movement speeds or high attack frequencies are associated with differences in prey choice and prey consumption [@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@toscanoPersonalityForagingBehavior2016]. Prey usually respond to predation cues by reducing their activity, or use rapid or unpredictable movements to successfully escape when they are chased [@Walker.etal2005;@Kelley.Magurran2011;@Herbert-Read.etal2017]. Such antipredator tactics result in reciprocal interactions where predators fine-tune hunting style to the type of prey that they encounter [@Estes.etal2003;@Woo.etal2008;@Phillips.etal2017;@Courbin.etal2018]. Winning the interaction would require that predators refine their tactics through periods of extensive practice [@DeLisle.etal2022;@Tinker.etal2009;@Dukas2019], suggesting that predators may learn through repeated interactions with their prey. This implies that a predator's prior experience and skills is an essential component that mediates hunting behaviour and the outcome of predator-prey interactions, with potential implications for community dynamics [@wooster2023].

Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. As predators practice and refine their tactics, they acquire expertise (e.g. skills, knowledge) that allows them to outperform novice hunters [@Dukas2019]. Through this process, it is hypothesized that individuals may specialize in a hunting tactic (e.g. always moving fast) if it allows them to maintain successful prey captures at each atempt [@Tinker.etal2009;@Dukas2019]. An alternative view is that learning operates as a mechanism driving behavioural flexibility. For instance, predators can gain expertise and information on their prey through repeated hunting atempts, and may thus hunt optimally by adjusting their behaviour to the type of prey that they encounter [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. However, theoretical and empirical evidence suggests that ecological interactions in early developmental stages can define how learning shapes the behaviour of individuals at later stages [@Stephens1993;@Ishii.Shimada2010;@vilapouca2021;@beyts2023], making it challenging to predict under which circumstances individuals should specialize in their foraging with experience.

When predators learn to hunt in environments where resources are predictable (or stable), individuals should specialize in their foraging behaviour through time because they need less energy and time to search for and handle prey [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015]. In contrast, while there are risks to attempt novel tactics by trial and error, predators can reduce the consequences of uncertainty when hunting in environments with fluctuating resources by learning to switch hunting tactics [@dall2010;@Szopa-Comley.Ioannou2022]. (*pourrait donner les exemples direct, et ensuite expliquer que cest toutefois tough d'établir les liens entre tout ça. Ultimement, la fitness doit appraitre quelque part.*)

Such contexts are important in defining which strategy (i.e. specialized vs flexible foraging)

il est toutefois difficile d'établir le niveau de fluctuation des resources lié à l'apprentissage et la stratégie de chasse.

Deux idées pour le paragraphe suivant :
1. Focus sur les contextes qui génèrent des différences, et qui devraient donc aussi mener à des différences de fitness (moins mettre l'accent sur le manque de consensus et plutôt focus sur les contextes, pour ensuite expliquer que donc ça devrait être mieux dans un que l'autre)

There is currently a lack of consensus on the fitness advantages of specialized vs flexible hunting strategies [@Phillips.etal2017]. Some studies have found greater success for individuals specializing on a hunting tactic [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others have found that individuals using flexible hunting tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021].

(*pourrait retirer en haut et commencer ici*)
many studies also show that both type of foragers can attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018].

finir le troisième paragraphe en disant que ultimement, si les individus vivent des degrés de fluctuations différentes, ils devraient pouvoir coexister dans la même population.
